---
title: "APPLIED PSET4"
author: "Aaliyah R. Robertson"
date: "May 14th, 2023"
output: pdf_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

<!-- .Rmd files use  markdown, a text mark up language, to provide formating.-->

<!--Text include within these strange arrows are comments and will not show up when you knit-->

# Front matter

This submission is my work alone and complies with the 30535 integrity policy.

Add your initials to indicate your agreement: \<\<\_\>\>

Upload your collaborators here: https://forms.gle/3mH1ofZZj9ZCgBHb8

Late coins used this pset: 0. Late coins left: 4. <!--You may use one for a given assignment.-->

# R for Data Science Exercises

<!--Note: Please do not restate questions in your submissions, directly write your answer--->

                 SECTION 4 DATA EXPLORATION IN JSON 
             
```{r}
setwd("/Users/aaliyahrobertson/Downloads")
waze_data <- read_rds("waze_data.rds")
```
4.1) Based on Chapter 24 and reviewing the waze data link, JSON is  hierarchical data structure that includes arrays and nested objects. JSON data can also support a wider range of data as opposed to the CSV type we have used in class. CSV data is in most cases, smaller and more compact compared to JSON data.

4.2)
```{r}
install.packages("jsonlite")
library(jsonlite)
library(tidyverse)
```
```{r}
json_string <- jsonlite::stream_in(url("https://www.waze.com/partnerhub-api/waze-feed-access-token/69c3e9b3-182c-4ec9-a5f3-c0e176568a3d?format=1"))

json_list <- as.list(json_string)
```
4.3) The length of the data is 7 and the names are "alerts, startTimeMillis, jams, endTimeMillis, irregularities, startTime, and endTime."

#What is the length 
```{r}
length(json_string)
```
#What are the names
```{r}
names(json_string)
```
#What items can be tibbles CHATGPT Query**
```{r}
str(json_string$alerts)
```
```{r}
str(json_list$irregularities)
```

                   SECTION FIVE DATA CLEANING
5.1) 
```{r}
library(testthat)
test_that("the waze data in the RDS format has the correct number of rows",
          {expect_equal(nrow(waze_data), 737357)})
```

5.2) Chicago Illinois
```{r}
cities_in_waze <- unique(waze_data$city)
print(cities_in_waze)
```
                           
5.3)
```{r}
library(dplyr)
library(tidyr)
waze_data <- separate(waze_data, col = "geo", into = c("latitude", "longitude"), sep = " ")
waze_data$latitude <- gsub("POINT|\\(|\\)", "", waze_data$latitude)
waze_data$longitude <- gsub("[()]", "", waze_data$longitude)
```

5.4) 
#ChatGPT query, 
The most commons alerts are weather hazards and traffic jams.
```{r}
library(ggplot2)
library(lubridate)

waze_data$hour <- hour(as.POSIXct(waze_data$ts))

ggplot(waze_data, aes(x = hour, fill = type)) +
  geom_histogram(binwidth = 1, position = "identity") +
  labs(x = "Hour of Day", y = "Number of Alerts", title = "Alerts by Hour of Day") +
  scale_fill_discrete(name = "Type of Alert")

```

5.5)
```{r}
library(lubridate)
waze_data$ts <- as.POSIXct(waze_data$ts, tz = "UTC")
waze_data$ts_ct <- with_tz(waze_data$ts, "America/Chicago")
```
```{r}
waze_data$ts_ct_rounded <- round_date(waze_data$ts_ct, "5 minutes")
```
                       
                        SECTION 7 WAZE SINGLE EVENT 

7.1a)****
#Define a bounding box around the event. Construct a data frame with all jams and accidents around the event.
```{r}
library(dplyr)
event_uuid <- "a42bc14b-e080-4621-9221-29dd86e553ce"
event_location <- waze_data %>%
  filter(uuid == event_uuid) %>%
  select(latitude, longitude) %>%
  distinct()
event_location
```

#ChatGPT query
```{r}
latitude <- as.numeric(event_location$latitude)
longitude <- as.numeric(event_location$longitude)

box_range <- 0.1  

latitude_min <- latitude - box_range
latitude_max <- latitude + box_range
longitude_min <- longitude - box_range
longitude_max <- longitude + box_range
```
```{r}
events_within_bounding_box <- waze_data %>%
  filter(as.numeric(latitude) >= latitude_min & as.numeric(latitude) <= latitude_max &
           as.numeric(longitude) >= longitude_min & as.numeric(longitude) <= longitude_max)
```

7.1b)
#What causes all these jams? Some googling might help.
Based off the variables and going through the data, it looks like the cause for these jams are within the subtype variable which varies. It includes  accidents, road construction, congestion, weather conditions, traffic incidents, etc, that are the reasons for traffic jams.

7.1c)
#Plot the number of jams 6AM-6PM CST.
```{r}
library(ggplot2)
uuid_plot <- events_within_bounding_box %>%
  filter(format(ts_ct, "%H:%M:%S", tz = "America/Chicago") >= "06:00:00" &
           format(ts_ct, "%H:%M:%S", tz = "America/Chicago") <= "18:00:00")
```
```{r}
jam_counts <- uuid_plot %>%
  mutate(hour = format(ts_ct, "%H")) %>%
  group_by(hour) %>%
  summarise(num_jams = n())
```
```{r}
ggplot(jam_counts, aes(x = hour, y = num_jams)) +
  geom_bar(stat = "identity", fill = "pink") +
  labs(title = "Number of Jams between 6AM-6PM CST", x = "Hour", y = "Number of Jams")
```



7.1d) 
#Next, propose a quantitative measure of traffic jam severity that combines the number of traffic
```{r}
severity_weights <- c(
  "JAM_STAND_STILL_TRAFFIC" = 5,
  "JAM_HEAVY_TRAFFIC" = 4,
  "ACCIDENT_MAJOR" = 5
)

severity_measure <- uuid_plot %>%
  mutate(severity_measure = confidence * severity_weights[subtype]) %>%
  summarise(total_severity = sum(severity_measure, na.rm = TRUE))

severity_measure$total_severity

```

7.1e)
#Plot this measure from 6AM-6PM CST.
```{r}
severity_plot <- uuid_plot %>%
  mutate(hour = as.numeric(format(ts_ct, "%H")),
         severity_measure = confidence * severity_weights[subtype]) %>%
  filter(hour >= 6 & hour <= 18) %>%
  group_by(hour) %>%
  summarise(total_severity = sum(severity_measure, na.rm = TRUE))

ggplot(severity_plot, aes(x = hour, y = total_severity)) +
  geom_line(color = "pink") +
  geom_point(color = "pink", size = 3) +
  labs(title = "Traffic Jam Severity between 6AM-6PM CST", x = "Hour", y = "Severity Measure")
```


                 


